\documentclass{article} % \documentclass{} is the first command in any LaTeX code.  It is used to define what kind of document you are creating such as an article or a book, and begins the document preamble

\usepackage{amsmath} % \usepackage is a command that allows you to add functionality to your LaTeX code

\title{Research} % Sets article title
\author{Sean Groenenboom \and Seger Sars \and Siem Vermeulen \and Angel Villanueva \and Ronan Vlak} % Sets authors name
\date{\today} % Sets date for date compiled
\newpage
\begin{document}

\maketitle % creates title using information in preamble (title, author, date)
\newpage

\tableofcontents % creates a table of contents
\newpage

\section{seger}
    \subsection{gamespeed}
    How do game engines handle game speed?
    \subsection{2d rendering}
    How do game engines render 2d graphics?
    \subsection{sprites}
    How do game engines handle sprites?
\newpage

\section{sean}
    \subsection{Multiplayer}
    Whhich are frequently used multiplayer structures?
    \subsection{input}
    How do game engines handle input?
    controller
    \subsection{ai}
    What kind of ai do game engines usually have?
    pathfinding?
\newpage

\section{siem}
    \subsection{audio}
    How is audio played on linux?
    multiple audio sources
    directional audio
    \subsection{save data}
    What are common save data structures?
    encryptie?
\newpage

\section{angel}
    \subsection{particles}
    What are particles in games?
    What are common methods game engines use to render particles?
    \subsection{physics}
    What are physics in games?

    common physics engines
    
\newpage

\section{ronan}
    \subsection{levels}
    What is a level in a video game?
    How are transitions between levels done?
    what are scenes?
    How are levels saved/stored?
    \subsection{gameobject}
    What is a gameobject in a game engine?
    How are components added to gameobjects?

\newpage

\section{temp}
    bullet hell engine
    \\
    2d engine.
    \\
    wow factor: veel entiteiten op het scherm. gamepseed ver kunnen verogen.
    performance tot de max.
\\
    seger:
    gamespeed                                   
    2d rendering 
    sprites (animatie + matrix berkeningen(rotere, schalen etc)) 
    \\
    sean:
    Multiplayer
    hoe connecten naar elkaar (ip adres, username?)
    wie host? Central? peer to peer?
    \\
    input + controller support
    ai
    \\
    siem:
    audio + adaptive music 
    data opslaan (voortgang, levels, statistieken, unlocks, achievements, etc)
    \\
    angel:
    particles?
    physics
    \\
    ronan:
    levels + levels switchen + scenes + hud
    gameobject
    
    
\newpage
\end{document} % This is the end of the document