\documentclass{article}

\title{Validation app design}
\author{Sean Groenenboom \and Seger Sars \and Siem Vermeulen \and Angel Villanueva \and Ronan Vlak} % Sets authors name
\date{\today} % Use \today for today's date or write a specific date, e.g., "October 31, 2023"

\begin{document}

\maketitle % Generates the title

\section{Introduction}
\begin{itemize}
    \item Validation app is a game.
    \item The game is a 2D bullet hell, inspired by Enter the Gungeon.
    \item The game is a rogue-lite. This means that when the player dies, some progress carries over into next runs.
\end{itemize}

\section{Core gameplay loop}
The game consists of three levels.
Each level contains five regular rooms, one shop room and a boss room.

\subsection{Player}
The player always starts a run with a standard amount of health and a starting weapon.

\subsection{Regular rooms}
Regular rooms contain only regular enemies.
The enemies are spawned semi-random: a random amount between set values, and at a random location in the room.
The type(s) of enemies spawned is predefined for each room.
Regular rooms contain pitfalls.

\subsection{Shop rooms}
Shop rooms contain a shopkeeper, which sells three randomly selected items to the player.
Each item has a standard fixed price.

\subsection{Boss rooms}
Boss rooms contain only a single boss enemy, and no other objects. 
When the boss is defeated. A magical door to the next level appears.
The boss then also drops 2 * the current level in EC. This is double if the level was executed without taking damage.
Boss rooms contain pitfalls.

\section{Gameplay mechanics}
- All bullets have infinite range (within the room).
\subsection{Enemies}
- Getting hit by an enemy cecreases the player health by 0.5 hearts.

\subsection{Currency}
Player upgrades can be unlocked using the EC currency (eternal credits). If you have enough EC you can unlock a new upgrade.
Upgrades cost a fixed amount, but differ between upgrades.
Items can be bought during a run using BS (Bullet Shells).

\subsection{Player items}
Items are bought during runs. The following types of items are available:
\begin{itemize}
    \item Health: increase health by 2 hearts. 15 EC to unlock.
    \item Damage: increase damage by 5 points. 30 EC to unlock.
    \item Speed: increase speed by 0.5 points. 15 EC to unlock.
    \item Fire rate: increase fire rate by 0.1 points. 15 EC to unlock.
    \item Bullet speed: increase bullet speed by 0.5 points. 15 EC to unlock.
    \item Buy weapon (shotgun): buy the shotgun weapon. 60 EC to unlock.
    \item Buy weapon (machine gun): buy the machine gun weapon. 60 EC to unlock.
\end{itemize}

Player upgrades mean unlocking new items, which can then be bought during a run.

\subsection{Environment}
\begin{itemize}
    \item Walls: the player cannot walk through walls. The edge of the level always consists of walls. Walls can be freely used within a room.
    \item Pitfalls: the player can fall into pitfalls. This decreases the player health by 1 heart. The player can always dodge over the pitfalls.
\end{itemize}

\section{Audiovisual design}
\begin{itemize}
    \item Sprites are stolen from Enter the Gungeon.
\end{itemize}

\section{Multiplayer}

\end{document}
