\section{Particle System}
\label{sec:particlesystem}
The particle system is a system that is used to create visual effects in the game.
\subsection{Particle Emitter}
The particle emitter is a component that can be used to create particle effects.
The particle emitter can be added to a GameObject, the emitter has a relative position, and it will spawn particles relative to the GameObject it is attached to.
The particle emitter has the following properties:
\begin{itemize}
    \item \texttt{emitter mode} - The mode of the emitter, can be set to Burst or Continuous.
    \item \texttt{particles} - The particles that make up the particle effect.
    \item \texttt{transform.x} - The x position of the particle effect relative to the GameObject it is attached to.
    \item \texttt{transform.y} - The y position of the particle effect relative to the GameObject it is attached to.
    \item \texttt{transform.rotation} - The rotation of the particle effect relative to the GameObject it is attached to.
    \item \texttt{angle} - in which direction the particles are emitted.
    \item \texttt{emissionRate} - The amount of particles emitted per second.
    \item \texttt{minimum particleLifetime} - The minimum lifetime of the particles.
    \item \texttt{maximum particleLifetime} - The maximum lifetime of the particles.
    \item \texttt{particleSpeed} - The speed of the particles.
    \item \texttt{acceleration} - The acceleration of the particles.
    \item \texttt{particleSize} - The size of the particles.
    \item \texttt{particleSizeShift} - The shift in size of the particles.
    \item \texttt{particleColor} - The color of the particles.
    \item \texttt{rotationSpeed} - The speed at which the particles rotate.
    \item \texttt{angularVelocity} - The angular velocity of the particles.
    \item \texttt{angularAcceleration} - The angular acceleration of the particles.
    \item \texttt{Color gradient} - The color gradient of the particles.
\end{itemize}

\subsection{Particle}
Particles are created by the particle emitter, and they make up the particle effect.
The particle emitter passes its own relevant properties to the particles.
The particles are then updated by the particle system, and they are rendered by the render system.
The particles have independent properties because they are updated and rendered independently.

The particle has the following properties:
\begin{itemize}
    \item \texttt{transform.x} - The x position of the particle.
    \item \texttt{transform.y} - The y position of the particle.
    \item \texttt{transform.rotation} - The rotation of the particle.
    \item \texttt{velocity} - The velocity of the particle.
    \item \texttt{acceleration} - The acceleration of the particle.
    \item \texttt{size} - The size of the particle.
    \item \texttt{angularVelocity} - The angular velocity of the particle.
    \item \texttt{angularAcceleration} - The angular acceleration of the particle.
    \item \texttt{lifetime} - The lifetime of the particle.
    \item \texttt{timeAlive} - The time the particle has been alive.
    \item \texttt{color gradient} - The color gradient of the particle.
\end{itemize}

The color gradient is a list of colors that the particle will switch between over its lifetime.
The particle can choose the nearest color in the gradient, or it can interpolate between the colors.

