\section{Render System}
\label{sec:rendersystem}
The render system is responsible for rendering the game.
The render system is called by the game loop, and it renders the current scene.
The render system goes through all the GameObjects in the scene, and it renders the Animation, ParticleEmitter, and Sprite components.
The render system also renders the UIObjects in the scene.

\subsection{Sprite}
The Sprite component is used to add a still image to a GameObject.

The Sprite Component has the following properties:
\begin{itemize}
    \item \texttt{texture} - The texture of the sprite.
    \item \texttt{width} - The width of the sprite.
    \item \texttt{height} - The height of the sprite.
    \item \texttt{flipx} - Whether the sprite is flipped horizontally.
    \item \texttt{flipy} - Whether the sprite is flipped vertically.
    \item \texttt{sprite source} - The source rectangle that defines which part of the texture is rendered(can be set to 0 to render all).
    \item \texttt{transform.x} - The x position of the texture relative to the GameObject it is attached to.
    \item \texttt{transform.y} - The y position of the texture relative to the GameObject it is attached to.
    \item \texttt{transform.rotation} - The rotation of the texture relative to the GameObject it is attached to.
\end{itemize}

\subsection{Animation}
The Animation component is used to add an animated image to a GameObject.
The Animation Component has the following properties:
\begin{itemize}
    \item \texttt{sprites} - The sprites that make up the animation.
    \item \texttt{transform.x} - The x position of the animation relative to the GameObject it is attached to.
    \item \texttt{transform.y} - The y position of the animation relative to the GameObject it is attached to.
    \item \texttt{transform.rotation} - The rotation of the animtion relative to the GameObject it is attached to.
    \item \texttt{flipX} - Whether the animation is flipped horizontally.
    \item \texttt{flipY} - Whether the animation is flipped vertically.
    \item \texttt{frameCount} - The amount of frames in the animation.
    \item \texttt{frameTime} - The time between each frame.
    \item \texttt{currentFrame} - The current frame of the animation.
    \item \texttt{loop} - Whether the animation should loop.
\end{itemize}
